---
title: "Predicting the present - 3.1 - Testing Author's Data"
output:
html_document:
df_print: paged
author: "Krushang Shah & Rajiv Basnet"
date: "6/23/2020"
---

# Predicting using the author data

## Setup
```{r setup, include=FALSE, message=FALSE}
library(tidyverse)
library(modelr)
library(plotly)
library(gtrendsR)
library(caret)

knitr::opts_chunk$set(echo = TRUE)
```

## Load Data
From the below mentioned CSV files, the appropriate one can be selected
```{r}
merged_author <- read_csv("merged_author.csv")
#merged <- read_csv("merged_extracted.csv")
#merged <- read_csv("merged_present.csv")
#merged <- read_csv("merged.csv")

merged_author$sales<-log(merged_author$sales)
merged_author$Period<-as.Date(merged_author$Period,"%m/%d/%Y")
rows <- nrow(merged_author)
```

## Model Summary
```{r}
model1_a <- lm(data = merged_author, sales~lag(sales, 1)+lag(sales,12))
summary(model1_a)
model2_a <- lm(data = merged_author, sales~lag(sales, 1)+lag(sales,12) + suvs + insurance)
summary(model2_a)
```

## Predict Current Month Data
```{r}
baseline_a <- merged_author
trends_a <- merged_author
# creating baseline 
for (i in 18:rows){
  merged_t_a <- merged_author[1:i-1,]
  model1_a <- lm(data = merged_t_a, sales~lag(sales, 1)+lag(sales,12))
  model2_a <- lm(data = merged_t_a, sales~lag(sales, 1)+lag(sales,12)+ suvs + insurance)
  baseline_a$sales[i] <- predict(model1_a,merged_author[1:i,])[i]
  trends_a$sales[i] <- predict(model2_a,merged_author[1:i,])[i]
}
```

## Compile Data
```{r}
actual_a <- merged_author[18:rows,] %>% mutate(label="actual")
baseline_a <- baseline_a[18:rows,] %>% mutate(label="baseline")
trends_a <- trends_a[18:rows,] %>% mutate(label="trends")
```

## Mean Absolute Error
```{r}
MAE(baseline_a$sales,actual_a$sales)
MAE(trends_a$sales,actual_a$sales)
```

## Recession MAE
```{r}
recession_trends_a <- trends_a %>% filter(Period>="2007-12-01"& Period<="2009-06-01")
recession_base_a <- baseline_a %>% filter(Period>="2007-12-01"& Period<="2009-06-01")
recession_actual_a <- actual_a %>% filter(Period>="2007-12-01"& Period<="2009-06-01")
MAE(recession_base_a$sales,recession_actual_a$sales)
MAE(recession_trends_a$sales,recession_actual_a$sales)
```

## Plot
```{r}
author_plot_data <- rbind(actual_a, baseline_a, trends_a)
ggplotly(ggplot(author_plot_data, aes(x=Period, y = sales, color = label, linetype = label))+
  geom_line()+
  scale_colour_manual(values=c("black", "red","grey"))+
  scale_linetype_manual(values = c("solid", "dashed", "solid"))+
  ylab('log(sales)')+
  xlab('period'))
```

# Predicting using the extracted data

## Load Data
From the below mentioned CSV files, the appropriate one can be selected
```{r}
#merged <- read_csv("merged_author.csv")
merged_extracted <- read_csv("merged_extracted.csv")
#merged <- read_csv("merged_present.csv")
#merged <- read_csv("merged.csv")

merged_extracted$sales<-log(merged_extracted$sales)
merged_extracted$Period<-as.Date(merged_extracted$Period,"%m/%d/%Y")
rows <- nrow(merged_extracted)
```

## Model Summary
```{r}
model1_e <- lm(data = merged_extracted, sales~lag(sales, 1)+lag(sales,12))
summary(model1_e)
model2_e <- lm(data = merged_extracted, sales~lag(sales, 1)+lag(sales,12) + suvs + insurance)
summary(model2_e)
```

## Predict Current Month Data
```{r}
baseline_e <- merged_extracted
trends_e <- merged_extracted
# creating baseline 
for (i in 18:rows){
  merged_t_e <- merged_extracted[1:i-1,]
  model1_e <- lm(data = merged_t_e, sales~lag(sales, 1)+lag(sales,12))
  model2_e <- lm(data = merged_t_e, sales~lag(sales, 1)+lag(sales,12)+ suvs + insurance)
  baseline_e$sales[i] <- predict(model1_e,merged_extracted[1:i,])[i]
  trends_e$sales[i] <- predict(model2_e,merged_extracted[1:i,])[i]
}
```

## Compile Data
```{r}
actual_e <- merged_extracted[18:rows,] %>% mutate(label="actual")
baseline_e <- baseline_e[18:rows,] %>% mutate(label="baseline")
trends_e <- trends_e[18:rows,] %>% mutate(label="trends")
```

## Mean Absolute Error
```{r}
MAE(baseline_e$sales,actual_e$sales)
MAE(trends_e$sales,actual_e$sales)
```

## Recession MAE
```{r}
recession_trends_e <- trends_e %>% filter(Period>="2007-12-01"& Period<="2009-06-01")
recession_base_e <- baseline_e %>% filter(Period>="2007-12-01"& Period<="2009-06-01")
recession_actual_e <- actual_e %>% filter(Period>="2007-12-01"& Period<="2009-06-01")
MAE(recession_base_e$sales,recession_actual_e$sales)
MAE(recession_trends_e$sales,recession_actual_e$sales)
```

## Plot
```{r}
extracted_plot_data <- rbind(actual_e, baseline_e, trends_e)
ggplotly(ggplot(extracted_plot_data, aes(x=Period, y = sales, color = label, linetype = label))+
  geom_line()+
  scale_colour_manual(values=c("black", "red","grey"))+
  scale_linetype_manual(values = c("solid", "dashed", "solid"))+
  ylab('log(sales)')+
  xlab('period'))
```