---
title: "new_replication_part2"
output: html_document
Authors: "Rajiv Basnet & Krushang Shah"
---


```{r, results='hide', message=FALSE}
library(tidyverse)
library(dplyr)
library(modelr)
library(scales)
library(lubridate)
library(stats)
```

```{r}
#read and clean unemployment_claims.csv

claims <- read.csv("unemployment_claims.csv")

# unemployment_claims.csv has been acquired and formatted from:
# https://oui.doleta.gov/unemploy/claims.asp

names(claims) <- c("date", "num_initial_claims")
claims <- claims %>% 
  mutate(num_initial_claims = as.numeric(gsub(",", "", num_initial_claims)), date = as.Date(date)) %>%  filter(date < "2020-05-30")
```


```{r}

#read and clean google trends data

jobs <- read.csv("jobs_google_trends.csv")
welfare <- read.csv("welfare_google_trends.csv")

welfare <- welfare %>% 
  mutate(date = as.Date(gsub("-", "/", paste(Month, "-01", sep = ""))))%>%
  select(-Month)

jobs <- jobs %>% 
  mutate(date = as.Date(gsub("-", "/", paste(Month, "-01", sep = "")))) %>%
  select(-Month)

```

NOTE: Google trends data are obtained on a monthly basis.
Thus, we will be matching the trends data with the claims data based on matching year and month. (Unlike the paper.)

```{r}

#merge google trends and unemployment data 

claims_new <- claims %>% mutate(year = year(date), month = month(date))

jobs_new <- jobs %>% mutate(year = year(date), month = month(date))
welfare_new <- welfare %>% mutate(year = year(date), month = month(date))

claims_with_trends <- claims_new %>% 
  inner_join(jobs_new, by = c("year", "month")) %>%
  select(-date.y) %>%
  rename(date = date.x) %>%
  inner_join(welfare_new, by = c("year", "month")) %>%
  select(-date.y) %>%
  rename(date = date.x)

claims_with_trends

```

```{r}
#our models

#date >= "2004-01-17" & date <= "2011-07-02"

claims_data <- claims_with_trends %>% 
  filter (date >= "2004-01-17" & date <= "2011-07-02") %>%
  mutate(log_num_claims = log(num_initial_claims))

model_baseline <- lm(log_num_claims ~ lag(log_num_claims, 1), data = claims_data)
model_with_trends <- lm(log_num_claims ~ lag(log_num_claims, 1) + jobs + welfare, data = claims_data)

summary(model_baseline)
summary(model_with_trends)
```

=> The baseline model's summary statistics here are pretty close to that mentioned in the paper.
=>  Likewise, the model with google trends also have statistics that are close to mentioned in the paper (not exact though).

Possible Causes: 
 - Use of monthly trends data to match the weekly data on original claims data.
 - Seasonal adjustment factors

Nevertheless, the addition of google trends data only had very slight improvement for the prediction (with regard to R-squared).


```{r}
# reproducing the visualizations

claims_data %>%
  filter (date >= "2008-01-01" & date <= "2012-01-01") %>%
  ggplot(aes(x = date, y = log_num_claims)) +
  geom_line()


```



